---
title: soundgen readme
output:
  html_document:
    variant: gfm
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include=FALSE}
stopifnot(require(knitr))
options(width = 90)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dev = "png",
  dpi = 150,
  fig.asp = 0.7,
  fig.width = 8,
  out.width = "80%",
  fig.align = "center"
)
library(soundgen)
```

[![CRAN Version](http://www.r-pkg.org/badges/version/soundgen)](https://cran.r-project.org/package=soundgen)
[![Downloads](http://cranlogs.r-pkg.org/badges/soundgen)](https://CRAN.R-project.org/package=soundgen)


R package for sound synthesis and acoustic analysis.   
Homepage with help, demos, etc: http://cogsci.se/soundgen.html   
Source code on github: https://github.com/tatters/soundgen   

Performs parametric synthesis of sounds with harmonic and noise components 
such as animal vocalizations or human voice. Also includes tools for spectral analysis, 
pitch tracking, audio segmentation, self-similarity matrices, morphing, etc.

# Key functions
* Sound synthesis from R console: `soundgen()`
* Shiny app for sound synthesis (opens in a browser): `soungen_app()`
* Acoustic analysis of a wav/mp3 file: `analyze()`
* Shiny app for editing intonation contours (opens in a browser): `pitch_app()`
* Measuring syllables, pauses, and bursts in a wav/mp3 file: `segment()`

For more information, please see the vignettes on sound synthesis and acoustic analysis:

`vignette("sound_generation", package="soundgen")`

`vignette("acoustic_analysis", package="soundgen")`   

Or, to open the vignettes in a browser:   
`RShowDoc('sound_generation', package = 'soundgen')`

`RShowDoc('acoustic_analysis', package = 'soundgen')`

# Example of sound synthesis
Use the `soundgen()` function to create a breathy moan:
```{r echo = FALSE, results = FALSE}
file.remove('man/figures/s.mp3')
```

```{r synthsesis, fig.width = 7, fig.height = 5}
s = soundgen(
  sylLen = 230,               # duration of voiced part, ms
  pitch = c(305, 280),        # pitch, Hz (goes down from 305 to 380 Hz)
  ampl = c(0, -20),           # amplitude, dB (gradual fade-out by 20 dB)
  rolloff = -30,              # strong f0, weak harmonics
  temperature = 0.05,         # some stochasticity in generation
  formants = c(260, 960, 1500, 2200,  # F1-F8 formant frequencies, Hz
               2600, 3600, 4200, 4500),
  noise = data.frame(
    time = c(-50, 120, 700),  # time of noise anchors
    value = c(-15, -5, -50)   # noise amplitude, dB
  ),
  rolloffNoise = 0,           # flat noise spectrum before adding formants
  addSilence = 0, samplingRate = 44100, pitchSamplingRate = 44100,
  play = TRUE, plot = TRUE, osc = TRUE, ylim = c(0, 6)
)
```

<p>
<audio controls style = "display: block"> <source src="man/figures/s.mp3" type="audio/mp3"> </audio>
</p>

```{r echo = FALSE, results = FALSE}
seewave::savewav(s, f = 44100, filename = 'man/figures/s.wav')
system('ffmpeg -y -i man/figures/s.wav man/figures/s.mp3')
file.remove('man/figures/s.wav')
```


# Example of acoustic analysis

Use the `analyze()` function to detect pitch and obtain other spectral descriptives of the sound we have just synthesized:
```{r analysis, fig.width = 7, fig.height = 5}
a = analyze(s, 44100, ylim = c(0, 6))
a[1:5, c('pitch', 'peakFreq', 'harmHeight', 'HNR', 'ampl', 'loudness')]
colnames(a)
```


# Installation
To install the current release from CRAN: `install.packages("soundgen")`

NB: Make sure all dependencies have been installed correctly! For problems with seewave, see http://rug.mnhn.fr/seewave/

On Macs, you may need to do the following:

* First install brew according to the instructions here: https://brew.sh/
* Then run the following from the terminal   
    `brew install libsndfile`   
    `brew install fftw`
* Finally, install soundgen in R:    
  `install.packages("soundgen")`

